---
title: "KIND community meetups"
execute: 
  freeze: false
listing:
  id: archive_list
  contents:
    - "commet/*.qmd"
  type: table
  categories: true
  fields: [title, date, categories]
---

These are our main weekly social learning activity. They're usually a slightly chaotic blend of chat, training session, tech demo, and show-and-tell.

+ [Download community meetup data as csv](data/KIND_community_meetups.csv)
+ [Download community meetup ideas as csv](data/KIND_community_meetups_new.csv)

:::: {.panel-tabset}

## Current community meetup schedule
```{r}
options(knitr.kable.NA = '')

readr::read_csv("data/KIND_community_meetups.csv") |>
  dplyr::mutate(Date = lubridate::dmy(Date)) |>
  dplyr::filter(Date >= lubridate::today() & !is.na(Date)) |>
  dplyr::select(-Suggestor) |>
  knitr::kable(caption = "Current community meetup schedule")
```

## Future community meetup ideas
```{r}
options(knitr.kable.NA = '')
readr::read_csv("data/KIND_community_meetups_new.csv") |>
  dplyr::select(!Date) |>
  knitr::kable(caption = "Future community meetup ideas")
```

## Archive of past sessions

::: {#archive_list}
:::


::::






